
\subsection{Changes from PR-ProvenanceDM-1.0-20181015}

\begin{itemize}
\item Rewording of the use cases in the Goal section.
\item Requirements divided into Model Requirements and Best Practices.
\item Restruturation of Section \ref{sec:datamodel}, removing the separation of the Core Model and the Extended Model. An overview is now given in \ref{sec:overview}, and each following subsection corresponds to a feature of the model, in relation with the goals. Each subsubsection is an element of the model presented in Figure \ref{fig:overview}.
\item \class{WasInformedBy} and \class{WasDerivedFrom} are optional and included in section \ref{sec:entity-activity-relations} on \class{Entity}-\class{Activity} relations. They are simple relations.
\item Addition of \class{ValueEntity} and \class{DatasetEntity} (with their description classes) as specific types of \class{Entity} classes (section \ref{sec:spec_entities}).
\item Grouping of the \class{Parameter}-\class{ParameterDescription} classes as an optional \class{ActivityConfiguration} package that also includes config files.
\item Modification of class attributes: 
\begin{itemize}
  \item \attribute{annotation} $\rightarrow$ \attribute{comment} (\attribute{annotation} $\rightarrow$ \attribute{description} in description classes)
  \item \attribute{creationTime} $\rightarrow$ \attribute{generatedAtTime} (W3C PROV term)
  \item \attribute{destructionTime} $\rightarrow$ \attribute{invalidatedAtTime} (W3C PROV term)
  \item added: \class{Agent}.\attribute{url}
  \item removed: \class{WasGeneratedBy}.\attribute{time}, \class{Entity}.\attribute{rights}, \class{Activity}.\attribute{status}, \class{ParameterDescription}.\attribute{xtype}, \class{ParameterDescription}.\attribute{arraysize}.
\end{itemize}
\item Datatype of a value is replaced by a valueType (see VO-DML document).
\item Only attributes are shown in tables, relations between classes are shown in diagrams.
\item Lists of terms are explicitely given for \class{Agent} roles, \class{ActivityDescription} types, \class{UsageDescription}/\class{GenerationDescription} types.
\item The section on serializations was removed (a dedicated document will be proposed). References to serialization in the text were removed, as well as related sentences, and Appendix A.
\item The section on accessing provenance information was removed (it was already reduced to a paragraph referencing external documents in preparation).
\item Appendix B on Links to other data models was removed.
\item Modeling Conventions and Data Types were added as appendices (taken from STC/Meas document).
\end{itemize}


\subsection{Changes from WD-ProvenanceDM-1.0-20180530}

\begin{itemize}
\item Separate core model (W3C only) and extended model (IVOA Provenance DM).
\item Add definitions for specialised entities, including all the description classes and parameters.
\item Add definitions for specialised relations.
\item Updated serialization of the description classes for web services.
\end{itemize}


\subsection{Changes from WD-ProvenanceDM-1.0-20170921}

\begin{itemize}
\item Moved ProvDAL (now ProvSAP) to separate document.
\item Moved ProvTAP section and full definition of VOTable serialisation to separate ProvTAP document.
\item Moved chapter 6 with use cases and ``How to use the data model'' to  separate Implementation Note \citep{std:ProvenanceImplementationNote}.
\item Moved section on links to other data model into appendix.
\item ParameterDescription: Added attributes \attribute{xtype} and \attribute{arraysize}.
\item Agent.roles: Removed ``PI'' alternative to ``coordinator''.
\item Use values of RightsType of DatasetDM, public, secure, proprietary, for \attribute{Entity.rights}.
\item Minor corrections in HiPS use case, 
%Appendix~\ref{sec:appendix-serialization-examples} 
and tables in TAP schema.
% and Appendix~\ref{sec:appendix-prov-votable} (TAP SCHEMA).
\item Minor correction in role names for hadStep/hadMember relationship.
\item Modified text on Parameters
\item Rename 3.4 section to Serialization of description classes for web services
\item Modified text on W3C serialization
\item add \attribute{location} and \attribute{value} attributes to \class{Entity}
\end{itemize}


\subsection{Changes from WD-ProvenanceDM-1.0-20161121}

\begin{itemize}
\item New appendix for PROV-VOTable/TAP SCHEMA tables added
\item Corrected and extended attribute tables and mapping tables for links with DatasetDM and SimDM.
\item Restructured Accessing provenance section by splitting it in two: Section for explaining the different serialization formats and differences to W3C serializations, Section ``Accessing provenance information''%~\ref{sec:provaccess} 
for describing the access protocols ProvDAL and ProvTAP.
\item Removed discussion section, since now all the topics are addressed in the main text.
\item Added paragraph on how to use the model in Section 6. %~\ref{sec:usecases-implementations}.
\item Shortened serialization examples, partially moved them to appendix.
\item Added paragraph on VOSI interface.
\item Added a proposed serialization of description classes.
\item Modified text on the content of EntityDescription, now seen as Entity attributes known before the Entity instance exists.
\item Renamed Section 6
%~\ref{sec:usecases-implementations} 
to stress that it explains applications of the model (use cases); implementation details and code examples can be found in Implementation Note \citep{std:ProvenanceImplementationNote}.
\item Complete rewrite of the ProvDAL section in Section ``Accessing provenance information'';%~\ref{sec:access_protocols};
new parameters, new figure and examples added.
\item Added additional figure for entity-activity relations.
\item Moved the figure showing relations between Provenance.Agent and Dataset.Party into the Section on data model links.
\item Extended the entity role examples in table \texttt{tab:entity-roles}.
\item Added links to provn and votable-serialization for HiPS-use case, added first part of provn as example in the HiPS-use case section.
\item More explanations on links to data models in a dedicated section, introduced subsections, added table with SimDM-mapping.
\item Moved detailed implementation section from appendix to a separate document (implementation note), shortened the use cases \& implementation section.
\item Attribute/class updates:
\begin{itemize}
    \item Added attribute \emph{votype} to \class{Activity}, can be used for ActivityFlows
    \item Added attribute \emph{time} to \class{Used} and \class{WasGeneratedBy}
    \item Added optional attributes \emph{Entity.creationTime} and \emph{EntityDescription.category}
    \item Added optional attributes \emph{Parameter.min},  \emph{Parameter.max},  \emph{Parameter.options}
    \item Removed the obscore/dataset attributes from EntityDescription, since they are specific for observations only and are not applicable to configuration entities etc.
    \item Use voprov:type and voprov:role in Table on Agent roles, i.e. replaced prov:person by Individual and prov:organization by Organization.
    \item Renamed \emph{label} attribute to \emph{name} everywhere, for more consistency with SimDM naming scheme (\emph{label} is reserved there for SKOS labels).
    \item Renamed attribute \emph{Entity.access} to \emph{Entity.rights} for more consistency with DatasetDM etc.
    \item Avoid double-meaning of \emph{description} (as reference and free-text description) by renaming the free-text description to \emph{annotation}. Mark description-references with arrows in attribute tables.
    \item Applied similar naming scheme to \emph{Parameter} and \emph{ParameterDescription}-classes
    \item Renamed \emph{docuLink} to \emph{doculink}
    \item Corrected attribute names in Table on mapping to DatasetDM.
\end{itemize}
\end{itemize}
