\documentclass[11pt,a4paper]{ivoa}
\input tthdefs

\usepackage[utf8]{inputenc}

\title{IVOA Provenance Data Model}

\ivoagroup{DM}

\author{Kristin Riebe}
\author{Michèle Sanguillon}
\author{Florian Rothmaier}
\author{Mathieu Servillat}
\author{Mireille Louys}
\author{Francois Bonnarel}

% \previousversion[????URL????]{????Funny Label????}
\previousversion[http://volute.g-vo.org/svn/trunk/projects/dm/provenance/description/ProvDM-0.1-20141008.pdf]{ProvDM-0.1-20141008.pdf}
       

% own definitions
\definecolor{todocolor}{rgb}{1,1,0.8}
\definecolor{darkred}{rgb}{0.6,0,0}
\definecolor{rose}{rgb}{1.0,0.88,0.88}
\definecolor{darkgrey}{rgb}{0.35,0.35,0.35}
%\newcommand{\TODO}[1]{%
%    \noindent%
%    \textcolor{todocolor}{\sffamily [\textbf{TODO:} #1]}%
%}

\newcommand{\TODO}[1]{%
    \noindent%
    \colorbox{todocolor}{%
            \parbox{0.85\linewidth}{\sffamily \textbf{TODO:}\\
            #1}
    }%    
    \vspace{2pt}

}

\newcommand{\Note}[1]{%
    \noindent%
    \textcolor{darkgrey}{{\sffamily Note:} \emph{#1}}%    
}


\newcommand{\paragraphlb}[1]{\paragraph{#1}\mbox{}\\} % paragraph with line break

\setlength{\fboxsep}{5pt}
%\setlength{\fboxrule}{1.5pt}
\newcommand{\warning}[1]{%
    \vspace{\baselineskip}
    \noindent
    \parbox{\linewidth}{%
        \colorbox{darkred}{%
            \parbox{0.7\linewidth}{\large \sffamily \textcolor{white}{Warning}}%
        }\\[-1pt]
        \noindent%
        \fcolorbox{darkred}{rose}{%
            \parbox{0.7\linewidth-2\fboxrule}{#1}%
        }%
    }%
    \vspace{\baselineskip}
}%

\begin{document}
\begin{abstract}
This draft is still an internal working draft, any comments welcome! 
The urls above also won't work yet, because this draft is only circulated 
via the volute-repository for now.

\warning{
    Everything described here is at a very 
    preliminary state and may change at any time!
}

\end{abstract}


\section*{Acknowledgments}

This document has been developed in part with support from the German
Astrophysical Virtual Observatory (BMBF Bewilligungsnummer 05A08VHA).

Thanks for fruitful discussions to (in alphabetical order):
Francois Bonnarel, Markus Demleitner, Jochen Klar, Gerard Lemson, 
Mireille Louys and Adrian Partl.



\section*{Conformance-related definitions}

The words ``MUST'', ``SHALL'', ``SHOULD'', ``MAY'', ``RECOMMENDED'', and
``OPTIONAL'' (in upper or lower case) used in this document are to be
interpreted as described in IETF standard, \citet{std:RFC2119}.

The \emph{Virtual Observatory (VO)} is
a general term for a collection of federated resources that can be used
to conduct astronomical research, education, and outreach.
The \href{http://www.ivoa.net}{International
Virtual Observatory Alliance (IVOA)} is a global
collaboration of separately funded projects to develop standards and
infrastructure that enable VO applications.


\section{Introduction}
\input{intro-general}
\input{intro-VOarchitecture}
\input{intro-requirements}
\input{intro-previousefforts}

\section{The provenance data model}
\input{datamodel-description}
\input{datamodel-discussion}


\section{Applications/Interactions with other Data models}
In this section we discuss how the provenance data model interacts with
other VO data models and how provenance information can be stored.

\TODO{Put this in appendix? Or in datamodel-description? Or even in introduction?}


\section{Accessing provenance information}
\input{provaccess}


\section{Use cases and implementations}\label{sec:usecases-implementations}
\input{usecases-implementations}


\appendix
\section{Changes from Previous Versions}
No official previous versions yet.
% No previous versions yet.  
% these would be subsections "Changes from v. WD-..."
% Use itemize environments.


\bibliography{ivoatex/ivoabib,prov-refs}


\end{document}
